Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-14 paper/paper.tex
to update the local in my office system with the remote

Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

This informal description is necessarily imprecise, with details differing widely between jurisdictions. For example,
the details of precisely which excess ballots are being transferred after candidates have been elected, the order in which surplus ballots are being transferred vary widely between different jurisdictions.

The goal of this paper is to advance formal methods in elections by demonstrating how STV vote counting can be verified in a modular, transparent, and practical feasible way over a minimal computational trust base:  we show how to synthesise efficient and provably correct verifiers for a large class of STV schemes in a modular and almost fully automated.

@@

We achieve this by first giving a formal definition of STV (called \emph{generic STV}) 
that  encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular STV vote counting scheme, but fixes both basic data structures and the mode of computation common to all instances of STV. Trustworthiness is achieved by
not verifying the \emph{software} that computes the result, but instead by verifying \emph{evidence} that substantiates the correctness of the count. Evidence records and corresponds to the individual steps taken in the computation of winners 
%(That is, our approach is an instance of verifiable computation \footnote{DP: reference}). 
%The evidence directly corresponds to the individual steps of the count
(and could so also be verified, or spot-checked manually), giving transparency. We minimize the computational trust base by verifying our executables down to the machine level using a combination of HOL \cite{hol} and CakeML \cite{cake}.

We observe that every instance of STV performs the same \emph{actions} (e.g. count first preferences, transfer a surplus, eliminate a candidate, etc.) where the precise interpretation of action differs. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  almost all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into machine-level verified executable using CakeML.

@@

In summary, we present a framework that allows us to formally specify instances of STV and sythesise an executable certificate verifier in a modular way that is itself verified down to the machine level.
\section{Evidences}
\label{sec:DataEv}

We are motivated to 
%engineer a trustworthy practical  framework to answer the 
tackle the following problem. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to  $\mathcal{A}$?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of (the source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. Therefore the question becomes what information is enough  for establishing transparency and verifiability of tallying  and what data structure for recording the information in the evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to attract in order to be elected, 
\item some candidates competing in the election, 
\item some number of vacancies for which candidates compete,
\item \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted. 
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action is constituted of  preconditions on when and how the action comes into effect, and postconditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example a constant precondition for applying elimination  asserts that there must be no uncounted ballot left and another one declares that no candidate should be electable at the current state of counting. A postcondition of elimination is that once a candidate is excluded from counting, the list of continuing candidates is updated accordingly so that the excluded candidate no longer received votes. Conjunctions of such declarations comprise what an action is and what an application of that action means.

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.  
Note that the states of computation recorded in all possible  evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  for $i\in\{1...j_{t}\}$, $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such precondition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in postconditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

We formally realise variations of STV algorithms by instantiation~(Definition~\ref{STVInst}) of the generic machine. Recall that the semantics of the machine~(Definition~\ref{STVMachine}) is formed by conjunctions of formally specified universally appearing pre- and postconditions in STV algorithms. An instantiation of the machine with an STV algorithm $\mathcal{A}$ happens by enriching the generic semantics by adding formal specifications of the legal clauses that are specific to $\mathcal{A}$. We hence come to define an operational semantics functioning in accordance with instructions of $\mathcal{A}$. 
\begin{definition}[Instantiation of the Machine]\label{STVInst}
Assume the generic machine $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$ as in Definition~\ref{STVMachine}. An instantiation $\hat{\mathcal{A}}$ for an STV algorithm $\mathcal{A}$ of  the generic STV machine $M$ is a triple $\langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$, where for each $t\in\mathcal{T}$, $S'_{t}$ $=$ $S_{t}\wedge\bigwedge_{\atop{i\leq j'_{t}}} \phi_{i}$. Each $\phi_{i}$ is the formal specification of a pre- or postcondition specific to $\mathcal{A}$.   
\end{definition}

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{The Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the framework is to produce means for validating concrete evidences. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.

@@

We formalise machine states as an inductive type  $\mathit{judgement}$ whose constructors are  $\mathit{NonFinal}$ and $\mathit{Final}$. 

@@

\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\;\HOLTyOp{tallies}\;\HOLTokenProd{}\;\HOLTyOp{piles}\;\HOLTokenProd{}\\\;\;\;\;\;\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\

@@

 The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $\mathit{((cand ~list)\times rat) list}$,   $\mathit{(cand\times rat)~list}$ and $\mathit{(cand\times(ballots~list)~list) ~list}$ where $\mathit{rat}$ is the HOL4 type of fractional numbers.

@@

 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a notion called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is distributed, according to next preferences shown on the ballots in the last parcel, to the continuing candidates.

@@

 Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need identifying the ballots constituting the last parcel of an elected. We hence formalise the pile to record, for each  continuing candidate, a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates. 
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of transfer-removed, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above.   
Implementing piles in this way enables us to tailor the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel or stepwise distribution of votes form eliminated candidates or the combination of both.

There is also a reason for choosing the type $rat$  
%as the underlying arithmetic for the formalisation, 
instead of e.g. floating numbers. Elections with an STV counting scheme have a small margin of victory especially for the last vacancy left to fill \cite{MBlo}. This margin may happen to be less than magnitude of the error caused by accumulation of rounding errors as it is with floating points.  Hence one must sensibly avoid the high  cost of electing a wrong candidate by falling into traps of imprecise calculations due to unintelligent choices. Using exact fractions allows safe handling of calculations. 
\subsection{The Semantics and its Auxiliary Components}
\label{sec:MachineSem}

The semantics of each machine transition label consists of conjunctions of formally specified general pre- and post-conditions across STV schemes which enforce when and how to take a tallying action and what the immediate effect is. To demonstrate the process of specification, implementation and verification of the machine transitions and their semantics, we discuss the transfer-elect transition.

>>>>

Our framework distils the similarities and differences of various instances of STV, giving a uniform and modular way of synthesising verifiers for its various instances. Our framework also provides a flexible and easy-to-use route for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using HOL4 and CakeML as the technical basis. First, we formally specify and verify the tools in HOL4. Then we obtain machine executable versions of the tools by relying on CakeML's verified proof translator and compiler. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate the feasibility of deploying the framework for verifying realistically sized elections run using an STV counting algorithm.
\end{abstract}

@@

This informal description is necessarily imprecise, with details differing widely between jurisdictions. For example,
the details of precisely which excess ballots are transferred after candidates have been elected, and the order in which surplus ballots are transferred vary widely between different jurisdictions.

The goal of this paper is to advance formal methods in elections by demonstrating how STV vote counting can be verified in a modular, transparent, and practical feasible way over a minimal computational trust base:  we show how to synthesise efficient and provably correct verifiers for a large class of STV schemes in a modular and almost fully automated way.

@@

We achieve this by first giving a formal definition of STV (called \emph{generic STV})
that  encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular STV vote counting scheme, but fixes both basic data structures and the mode of computation common to all instances of STV. Trustworthiness is achieved by
not verifying the \emph{software} that computes the result, but instead by verifying \emph{evidence} that substantiates the correctness of the count. Evidence records and corresponds to the individual steps taken in the computation of winners
%(That is, our approach is an instance of verifiable computation \footnote{DP: reference}).
%The evidence directly corresponds to the individual steps of the count
(and so could also be verified, or spot-checked manually), giving transparency. We minimize the computational trust base by verifying our executables down to the machine code level using a combination of HOL~\cite{hol} and CakeML~\cite{cake}.

We observe that every instance of STV performs the same \emph{actions} (e.g. counting first preferences, transferring surplus votes, eliminating candidates, etc.) where the precise interpretations of those actions differ. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of a count, then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and post-conditions for each rule, as well as a computable predicate to decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  almost all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into a machine-level verified executable using CakeML.

@@

In summary, we present a framework that allows us to formally specify instances of STV and sythesise an executable certificate verifier in a modular way that is itself verified down to the machine level.
\section{Evidence}
\label{sec:DataEv}

We are motivated to
%engineer a trustworthy practical  framework to answer the
tackle the following problem: how can one verify that any execution of any implementation $\mathcal{P}$ of an arbitrary  known STV algorithm $\mathcal{A}$ (whose source code may be secret) correctly computes the end result according to $\mathcal{A}$?

To begin,
%with addressing the problem,
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of (the source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

Therefore, the evidence must be sufficiently informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per the counting scheme. In this way, we can be sure that winners truly reflect will of the voters. We must determine what information is enough to establish transparency and verifiability of tallying, and what data structure for recording the information in the evidence should be chosen.
\subsection{Characterisation of Evidence}
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows:
%Each STV scheme operates on the following data.
\begin{itemize}
\item \emph{ballots} recording the voters' votes. Each ballot is a pair of a list expressing the voter's preferences, and a fractional transfer value,
\item a \emph{quota} which is the least amount of votes needed to attract in order to be elected,
\item the \emph{candidates} competing in the election,
\item the \emph{number of vacancies} for which candidates compete,
\item \emph{discrete states of computation} encapsulating the information needed to determine how the tallying has progressed from the beginning to the end, and to be able to independently verify its correctness.
\end{itemize}
There are two kinds of computation states. One kind are final states, where tallying has reached a halting state by finding the winners of the election. The other kind consists of intermediates states where data is manipulated according to the counting mechanism of the STV scheme used, to eventually obtain an end result. Each intermediate state comprises seven pieces of data:
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate,
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election.
 \end{itemize}
The non-final and final states specified above provide us with all the necessary information.
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given, and that $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$.
Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. Let $\hat{\omega}$ of the execution be the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation, with the initial state containing $x$, the input, and the final state giving $y$ as the algorithm's output.
\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics to describe the way this data type is manipulated as part of the process of  evidence validation.
On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand, it must also facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings.
%\begin{figure}
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates,
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates,
\item\textbf{eliminate} to exclude one or some candidates,
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates,
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left,
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies.
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action consists of pre-conditions on when and how the action comes into effect, and post-conditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example, two pre-conditions for applying elimination asserts that (i) there must be no uncounted ballots left, and (ii) that no candidate should be electable at the current state of counting. A post-condition of elimination is that once a candidate is excluded from counting, the list of continuing candidates is updated accordingly so that the excluded candidate no longer receives votes. Conjunctions of such declarations comprise what an action is and what an application of that action means.

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Finally, the pre- and post-conditions which define the actions result in a small-step operational semantics for the machine.
Note that the states of computation recorded in evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  for $i\in\{1...j_{t}\}$, $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} of the machine.
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying \textbf{transfer-removed} action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of \textbf{transfer-removed}. There are also variations in post-conditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

We formally realise variations of STV algorithms by instantiation~(Definition~\ref{STVInst}) of the generic machine. Recall that the semantics of the machine~(Definition~\ref{STVMachine}) is formed by conjunctions of formally specified universally appearing pre- and post-conditions in STV algorithms. An instantiation of the machine with an STV algorithm $\mathcal{A}$ happens by enriching the generic semantics by adding formal specifications of the legal clauses that are specific to $\mathcal{A}$. We hence come to define an operational semantics functioning in accordance with instructions of $\mathcal{A}$.
\begin{definition}[Instantiation of the Machine]\label{STVInst}
Assume the generic machine $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$ as in Definition~\ref{STVMachine}. An instantiation $\hat{\mathcal{A}}$ for an STV algorithm $\mathcal{A}$ of  the generic STV machine $M$ is a triple $\langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$, where for each $t\in\mathcal{T}$, $S'_{t}$ $=$ $S_{t}\wedge\bigwedge_{\atop{i\leq j'_{t}}} \phi_{i}$. Each $\phi_{i}$ is the formal specification of a pre- or post-condition specific to $\mathcal{A}$.
\end{definition}

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}}
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.
\section{The Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.
\paragraph*{Implementation} Our final objective in creating the framework is to validate concrete evidence. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define executable functions in the implementation sub-module to computationally realise the specification module's logical declarations. In particular, we implement decision procedures whose computational content provably realises the specification of the machine semantics.
\paragraph*{Verification} Here we formally prove that the implementations meet their respective specifications.
We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack the skills in the functional programming style, which we rely on for implementing verifiers in HOL4. However, the same user may well have had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, allows users to work at the more abstract specification level.

@@

We formalise machine states as an type  $\mathit{judgement}$, with two constructors $\mathit{NonFinal}$ and $\mathit{Final}$.

@@

\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{tallies}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{piles}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\

@@

 The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $((\mathit{cand}~\mathit{list})\times\mathit{rat})~\mathit{list}$,   $(\mathit{cand}\times\mathit{rat})~\mathit{list}$ and $(\mathit{cand}\times(\mathit{ballots}~\mathit{list})~\mathit{list})~\mathit{list}$ where $\mathit{rat}$ is the HOL4 type of the rational numbers.

@@

 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $(\mathit{cand}\times(\mathit{ballots}~\mathit{list}))~\mathit{list}$;
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a notion called ``last parcel''. The last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is distributed, according to next preferences shown on the ballots in the last parcel, to the continuing candidates.

@@

 Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need to identify the ballots constituting the last parcel of an elected candidate. We hence formalize the pile to record, for each  continuing candidate, as a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates.
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of \textbf{transfer-removed}, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above.
Implementing piles in this way enables us to tailor the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel or stepwise distribution of votes form eliminated candidates or a combination of both.

There is also a reason for choosing the type $rat$
%as the underlying arithmetic for the formalisation,
instead of e.g., floating numbers. Elections with an STV counting scheme have a small margin of victory especially for the last vacancy left to fill~\cite{MBlo}. This margin may happen to be less than magnitude of the error caused by accumulation of floating point number rounding errors.  Hence one must sensibly avoid the high  cost of electing a wrong candidate by falling into traps of imprecise calculations due to unintelligent choices. Using exact fractions allows safe handling of calculations.
\subsection{The Semantics and its Auxiliary Components}
\label{sec:MachineSem}

The semantics of each machine transition label consists of conjunctions of formally specified general pre- and post-conditions across STV schemes which enforce when and how to take a tallying action and what the immediate effect is. To demonstrate the process of specification, implementation and verification of the machine transitions and their semantics, we discuss the \textbf{transfer-elect} transition.
